Supplementary Information: Vulnerability of tree growth

### 2025-02-21 doc/manuscript.Rmd
manuscript edits to statistical models

To further disentange the drivers of stem-level drought sensitivity, we modelled:

$$
sensitivity \sim DBH + CII + TWI + (DBH+CII+TWI | Species)
$$

We chose to model a species random effect on all slopes here to account for variation across species instead of specific traits which may have different effects with each of the variables.

All statistical analyses were performed using R version 4.4.0 [@rcoreteam_language_2024]. 
Mixed models were fit through Bayesian regressions using MCMC methods using the package _brms_ [REF].
We ran models using gaussian priors for the data distribution.
We ran 4 MCMC chains for 3000 iterations with 1000 of these set as warmup.
We extracted draws, calculated distributions using functions from the _posterior_ package [REF].

@@

@@

![***Figure 4: Drivers of variation in sensitivity among individuals***](display/Fig4.png)

>>>>

To further disentange the drivers of stem-level drought sensitivity, we simultaneously modelled:

$$
sensitivity \sim DBH + CII + TWI + (DBH+CII+TWI | Species)
$$ 
and 
$$
CII \sim DBH
$$

Simultaneously modelling the effect of variables on sensitivity and the effect of DBH on CII in this way allows us to disentangle causual linkages between these processes. 
We chose to model a species random effect on all slopes here to account for variation across species instead of specific traits which may have different effects with each of the variables. 
However, we also ran simpler models with species random effects only on the intercept, the results of which are reported in the Supplementary Material. 
Across all models, we used Gaussian priors for the distribution of the response variable sensitivity, considered CII as an monotonic predictor (ordered factor) and centred and scaled the other predictors. 
We modelled CII as an ordinal categorical variable with a cumulative logit link function. 
In the combined form, the model therefore used a multivariate distribution of Gaussian and logit. 
Finally, to test full mediation effects, or the hypothesis that DBH effects on sensitivity are fully explained by its effects on exposure, we tested the same models, removing the DBH as a predictor of sensitivity. 
We then compared the fits of the partial mediation and full mediation models using Leave-One-Out Cross Validation.

All statistical analyses were performed using R version 4.4.0 [@rcoreteam_language_2024]. 
Mixed models were fit through Bayesian regressions using MCMC methods using the package _brms_ [REF].
We ran 4 MCMC chains for 3000 iterations with 1000 of these set as warmup.
We extracted draws, calculated distributions using functions from the _posterior_ package [REF].

@@

![***Figure 4: Drivers of variation in sensitivity among individuals***](display/Fig4_alternate.png)

@@

![***Figure 5: Effects of size, exposure and water availability***](display/Fig5.png)
